---
title: "What overlap is there (if any) with trans eQTLs?"
author: "Britta Velten"
date:  "`r format(Sys.Date(),'%e %B, %Y')`"
output: BiocStyle::html_document
params:
  section_name : "10f_trans_eQTL_overlap"
  subsection_name : "10f_01_trans_eQTL_overlap_overall_signal"

---

# Preparations
```{r prep, warning=FALSE, message=FALSE, echo = F}
# to render pngs in html
library(Cairo)
knitr::opts_chunk$set(fig.path = plotsdir, dev=c("CairoPNG", "pdf"), echo = F, warning = F, message = F)
```

# Comparison of Effect Sizes

```{r}
ggplot(control_cis_effects, aes(x = cis_eqtl_beta, y = control_beta, col = -log10(control_p_value))) + 
  geom_point(size = 0.4) + theme_bw()

```


# Down-regulation of Target Gene

```{r}
### down-regulation of target
df4plot <- left_join(trans_eQTL_df, expression_variance)
ggplot(df4plot, aes(x = abs(control_beta), y = abs(target_lfc), label = gene_name, col = sc_var)) + 
  ggrepel::geom_text_repel(size = 8/.pt) + 
  xlab("Natural effects") + ylab("CRISPR down-regulation") + 
  geom_point(size = 0.4) + geom_abline(slope = c(1,2), lty = 3, col = 'red') + theme_bw()
```
## Top Effects

Consider the SNPs where there is significant on-target expression and also signal for natural expression variation in our control lines.

```{r}
df <- left_join(left_join(trans_eQTL_df, expression_variance) %>%
  mutate(crispr_n_trans = target_meta$n_downstream_excl_target[match(gene_name, target_meta$gene)]),
  effect_cor_df %>% dplyr::select(c("snp_id", 'beta_cor', 'pval_cor')))
snps2consider <- df %>%
  dplyr::filter(pval_adj < sig_pval_thresh & crispr_n_trans > 1) %>%
  mutate(ratio_of_effect = abs(control_beta/target_lfc)) %>%
  slice_max(ratio_of_effect, n = 20) 
```


## Effects with Signal

Figure out which cis eQTLs have some signal:

```{r, fig.width = 4, fig.height = 4}
s <- '11_805592_C_G'
df2plot <- cis_eQTL_meta %>%
    dplyr::filter(snp_id == s ) 
c <- cor(df2plot$trans_eQTL_beta, df2plot$crispr_lfc)
p <- ggplot(df2plot, aes(x = trans_eQTL_beta, y = crispr_lfc,
                      col = ifelse(crispr_pval_adj < 0.1, 'y', 'n'),
                      label = ifelse(trans_gene_name == cis_gene_name, cis_gene_name, ''))) + 
    ggrepel::geom_text_repel() + 
    scale_color_manual('', values = c('y' = 'red', 'n' = 'gray')) + 
    ggtitle(paste0(df2plot$cis_gene_name, " - ", df2plot$snp_id)) + 
    geom_point()+ 
    annotate('text', label = paste0("Pearson's R = ", formatC(c, 4)), x = -Inf, y = -Inf, hjust = -0.15, vjust = -2) +
    theme_bw() + theme(legend.position = 'none') 
p
  

```


Do the high effect genes have smaller effect sizes?

```{r}
df4plot <- df %>%
  dplyr::filter(pval_adj < sig_pval_thresh) %>%
  mutate(effect_size_ratio = control_beta/target_lfc)
ggplot(df4plot, aes(x = abs(effect_size_ratio), y= crispr_n_trans)) + 
  geom_point()
```




# Distribution of Correlation

Do the eQTLs that correlate well have strong cis eQTL effects?

```{r}
df4plot <- left_join(control_cis_effects, effect_cor_df)
ggplot(df4plot, aes(x = cis_eqtl_beta, y = beta_cor)) + 
  geom_point()
```
# Reproducibility in Strong Effects

```{r}

```

# SessionInfo
```{r}
sessionInfo()
```






