---
title: "What overlap is there (if any) with trans eQTLs?"
author: "Britta Velten"
date:  "`r format(Sys.Date(),'%e %B, %Y')`"
output: BiocStyle::html_document
params:
  section_name : "10f_trans_eQTL_overlap"
  subsection_name : "10f_01_trans_eQTL_overlap_overall_signal"

---

# Preparations
```{r prep, warning=FALSE, message=FALSE, echo = F}
# to render pngs in html
library(Cairo)
knitr::opts_chunk$set(fig.path = plotsdir, dev=c("CairoPNG", "pdf"), echo = F, warning = F, message = F)
```

# Power of Tests

How many targets have a cis-eQTL? (What is the distribution of # of cis-eQTLs?)

How many expressed genes have a trans eQTL? How many expressed genes were tested?


# Overall Signal

## Comparison of p-values

```{r}
cor(df$trans_eQTL_p_value, df$crispr_pval)
```

```{r}
ggplot(df, aes(x = trans_eQTL_p_value, y = crispr_pval)) + 
  geom_point() + theme_bw()
```

Subset this to only the significant CRISPR effects:

```{r}
ggplot(df %>%dplyr::filter(crispr_pval_adj < 0.1), aes(x = trans_eQTL_p_value, y = crispr_pval)) + 
  geom_point() + theme_bw()
```

```{r}
df4qq_plot <- df %>% 
  mutate(status = ifelse(crispr_pval_adj < 0.1, 'crispr_sig', 'not_sig'))

ggplot(df4qq_plot, 
                         aes(sample = trans_eQTL_p_value, col = status)) + 
  stat_qq(distribution = stats::qunif) 

qq_plot_coords_sig <- ggplot(df4qq_plot %>%
                           dplyr::filter(status == 'crispr_sig'), 
                         aes(sample = trans_eQTL_p_value, col = status)) + 
  stat_qq(distribution = stats::qunif) 

qq_plot_coords <- ggplot_build(qq_plot_coords)$data[[1]]
  mutate(status = 'is_pair')
  
eQTL_p_vals <- df %>%
  dplyr::filter(crispr_pval_adj < 0.1) %>%
  mutate(status = 'is_pair')

background_lfcs <- df %>%
  dplyr::filter(crispr_pval_adj > 0.1) %>%
  mutate(status = 'not_pair')
set.seed(0)
background_lfcs <- background_lfcs[sample(1:dim(background_lfcs)[1], size = dim(eQTL_p_vals)[1]),]

trans_eQTL_coords <- ggplot(eQTL_p_vals, aes(sample = trans_eQTL_p_value)) + 
  stat_qq(distribution = stats::qunif)
trans_eQTL_coords <- left_join(
  eQTL_p_vals %>% dplyr::select(c('cis_gene_name', 'trans_gene_name', 'pval_lm' = 'trans_eQTL_p_value')),
  ggplot_build(trans_eQTL_coords)$data[[1]] %>% dplyr::select(c('pval_lm' = "sample", "theoretical"))
) %>%
  mutate(status = 'crispr_sig')
background_coords <- ggplot(background_lfcs, aes(sample = trans_eQTL_p_value)) + 
  stat_qq(distribution = stats::qunif)
background_coords <- left_join(
  background_lfcs %>% dplyr::select(c('cis_gene_name', 'trans_gene_name', 'pval_lm' = 'trans_eQTL_p_value')),
  ggplot_build(background_coords)$data[[1]] %>% dplyr::select(c('pval_lm' = "sample", "theoretical"))
) %>%
  mutate(status = 'not_crispr_sig')


df4plot <- as.data.frame(bind_rows(
  background_coords,
 trans_eQTL_coords
))


ggplot(df4plot, aes(x = -log10(theoretical), y = -log10(pval_lm), col = status)) + 
  #ggrepel::geom_text_repel(size = 3) + 
  scale_color_manual('', values = c("crispr_sig" = target_downstream_col, 'not_crispr_sig' = 'gray')) + 
  scale_x_log10() + scale_y_log10()+
  xlab("Theoretical (Log10 p-value)") + ylab("Sample  (Log10 p-value)") + 
  geom_point(size = 0.4) + geom_abline(slope = 1, lty = 3, col = 'black') + theme_bw() + theme(aspect.ratio = 1)

```

## Correlation of Effect Sizes

```{r}
ggplot(df, aes(x = trans_eQTL_beta, y = crispr_lfc*snp_beta)) + 
  geom_point() + theme_bw()
```

# Per SNP

## Lowest correlations

```{r fig.width = 12, fig.height = 8}

snps2plot <- trans_effect_info %>%
  dplyr::filter(beta_cor < -0.15)
plots <- lapply(snps2plot$snp_id, FUN = function(s){
  df2plot <- df %>%
    dplyr::filter(snp_id == s )
  c <- cor(df2plot$trans_eQTL_beta, df2plot$crispr_lfc)
  p <- ggplot(df2plot, aes(x = trans_eQTL_beta, y = crispr_lfc,
                      col = ifelse(crispr_pval_adj < 0.1, 'y', 'n'),
                      label = ifelse(trans_gene_name %in% trans_genes, trans_gene_name, ''))) + 
    ggrepel::geom_text_repel() + 
    scale_color_manual('', values = c('y' = 'red', 'n' = 'gray')) + 
    ggtitle(paste0(df2plot$cis_gene_name, " - ", df2plot$snp_id)) + 
    geom_point()+ 
    annotate('text', label = paste0("Pearson's R = ", formatC(c, 4)), x = -Inf, y = -Inf, hjust = -0.15, vjust = -2) +
    theme_bw() + theme(legend.position = 'none') 
  #p
  return(p)

})

ggarrange(plotlist = plots)

```

## Highest Correlations


```{r fig.width = 12, fig.height = 8}

snps2plot <- trans_effect_info %>%
  slice_max(beta_cor, n = 6)
plots <- lapply(snps2plot$snp_id, FUN = function(s){
  df2plot <- df %>%
    dplyr::filter(snp_id == s )
  c <- cor(df2plot$trans_eQTL_beta, df2plot$crispr_lfc)
  p <- ggplot(df2plot, aes(x = trans_eQTL_beta, y = crispr_lfc,
                      col = ifelse(crispr_pval_adj < 0.1, 'y', 'n'),
                      label = ifelse(trans_gene_name %in% trans_genes, trans_gene_name, ''))) + 
    ggrepel::geom_text_repel() + 
    scale_color_manual('', values = c('y' = 'red', 'n' = 'gray')) + 
    ggtitle(paste0(df2plot$cis_gene_name, " - ", df2plot$snp_id)) + 
    geom_point()+ 
    annotate('text', label = paste0("Pearson's R = ", formatC(c, 4)), x = Inf, y = -Inf, hjust = 1.25, vjust = -2) +
    theme_bw() + theme(legend.position = 'none') 
  #p
  return(p)

})

ggarrange(plotlist = plots)

```


# SessionInfo
```{r}
sessionInfo()
```






